# Analysis

```{r init}
#| label: Analysis

source("globals.R")
```

## Data Overview

### **Expected:**

-   fig/text: Data coverage and share of “in operation” deals in the total deals.

-   xxxx unique investors identified, around xxxx% of investors registered in tax havens.

-   fig: Temporal distribution of land deals

### Data availability

```{r}
#| label: check-availability

# check data.csv 
data_availability(
  data_name = "data",
  fields = c("initiation_year", "current_intention_of_investment", "investment_type", "target_country_code_alpha2","world_region","country_iso_code")
)

# check desag.csv
data_availability(
  data_name = "desag",
  fields = c("role", "country_iso2", "target_country_code_alpha2", "is_tax_haven", "current_implementation_status","initiation_year")
)

```

```{r}
#| label: value-summary

value_summary(data, "current_implementation_status")
value_summary(data, "world_region")

value_summary(desag, "is_tax_haven")
value_summary(desag, "role")

value_summary(investors, "from_tax_haven")
```

### How many deals involve tax havens?

```{r}
#| label: deals-involve-TaxHaven

deal_taxhaven <- desag |>
  group_by(deal_id) |>
  summarise(
    num_tax = sum(is_tax_haven, na.rm = TRUE),
    any_tax_haven = num_tax > 0,
    current_implementation_status = first(current_implementation_status),
    initiation_year = first(initiation_year),
    .groups = "drop"
  )

value_summary(deal_taxhaven, "any_tax_haven")
value_summary(deal_taxhaven, "num_tax")

deal_taxhaven |>
  filter(current_implementation_status == "In operation") |>
  summarise(
    total_in_operation = n(),
    tax_haven_deals    = sum(any_tax_haven),
    share              = scales::percent(tax_haven_deals / total_in_operation, accuracy = 0.1)
  ) |>
  knitr::kable(
    caption = "In operation deals involving tax haven investors",
    align = "c"
  ) |>
  kableExtra::kable_styling(full_width = FALSE, position = "center")


```

To quantify the involvement of tax-haven–registered investors, we aggregated the investor-level information in *desag_data* to the deal level. **A deal is considered to involve a tax haven if at least one investor within its ownership chain is registered in a jurisdiction listed as a tax haven.**

The results show that **1318** deals (representing **24.7%** of all deals) involve at least one tax-haven investor. This indicates that tax-haven jurisdictions play a non-negligible role in land investment structures, either through shareholders, controlling shareholders, or ultimate owners.

When considering only deals currently ***in operation***, **946** deals out of **3446** are associated with tax-haven investors, corresponding to **27.5%**.

### How many unique investors involve tax havens?

```{r}
#| label: investor-involve-TaxHaven

investor_taxhaven <- desag |>
  group_by(bvd_id_number) |>
  summarise(
    investor_country = paste(unique(country_iso2), collapse = ", "),
    roles = paste(unique(role), collapse = ", "),
    num_deals = n_distinct(deal_id),
    is_tax_haven_investor = any(is_tax_haven, na.rm = TRUE),
    .groups = "drop"
  )

value_summary(investor_taxhaven, "is_tax_haven_investor")
```

Across the entire database, we identified **12,405 unique investors** involved in at least one land deal. Among them, **898 investors** are registered in jurisdictions classified as tax havens, representing **7.2%** of all unique investors.

```{r}
#| label: investor-from-TaxHaven

# fliter the investor registered in tax haven
taxhaven_roles <- investor_taxhaven |>
  filter(is_tax_haven_investor == TRUE) |>
  tidyr::separate_rows(roles, sep = ",") |>
  mutate(roles = trimws(roles))

value_summary(taxhaven_roles, "roles")
```

### How does the number of land deals changed over time?

```{r}
#| label: fig-temporal-distribution
#| fig-cap: "Temporal distribution of total land deals"

data |> 
  filter(!is.na(initiation_year)) |> 
  count(initiation_year, name = "n_deals") |> 
  ggplot(aes(x = initiation_year, y = n_deals)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    x = "Initiation year",
    y = "Number of deals",
    title = "Temporal distribution of land deals"
  ) +
  theme_minimal()
```

```{r}
#| label: fig-temporal-distribution-operation
#| fig-cap: "Temporal distribution of land deals in operation"

data |> 
  filter(
    initiation_year >= 2000,
    current_implementation_status == "In operation"
  ) |> 
  count(initiation_year, name = "n_deals") |> 
  ggplot(aes(x = initiation_year, y = n_deals)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(
    x = "Initiation year",
    y = "Number of deals",
    title = "Temporal distribution of land deals in operation"
  ) +
  theme_minimal()

# In-operation deals since 2000: Total vs Tax-haven involved
deal_taxhaven |> 
  filter(
    initiation_year >= 2000,
    current_implementation_status == "In operation"
  ) |> 
  group_by(initiation_year) |> 
  summarise(
    total = n(),
    taxhaven = sum(any_tax_haven),
    .groups = "drop"
  ) |> 
  ggplot(aes(x = initiation_year)) +
  geom_line(aes(y = total, color = "Total deals"), linewidth = 1) +
  geom_point(aes(y = total, color = "Total deals"), size = 2) +
  geom_line(aes(y = taxhaven, color = "Tax-haven deals"), linewidth = 1) +
  geom_point(aes(y = taxhaven, color = "Tax-haven deals"), size = 2) +
  scale_color_manual(
    values = c("Total deals" = "steelblue", "Tax-haven deals" = "red")
  ) +
  labs(
    x = "Initiation year",
    y = "Number of deals",
    color = NULL,
    title = "In-operation deals since 2000: Total vs Tax-haven involved"
  ) +
  theme_minimal()

```

## Distribution Patterns

1.  bar: Share of tax haven vs. non–tax haven investments

2.  line: Evolution of tax haven investments over time -\> Analysis of whether the use of tax havens has increased over time

3.  bar: Top 10 tax haven investor countries -\> Identify the primary source countries of tax havens

4.  bar: Top 10 target countries for tax haven investors -\> Identify the principal countries of investment

5.  bar: Tax haven investments by sector -\> Comparing the proportion of tax havens across different investment sectors

6.  Map: Global investment flows from tax havens to target countries -\> Visualising the global investment pathways of tax havens

### Evolution of tax haven investments over time

```{r}
#| label: fig-temporal-distribution-taxhaven
#| fig-cap: "Temporal distribution of land deals in operation"

deal_taxhaven |> 
  filter(initiation_year >= 2000) |> 
  count(initiation_year, any_tax_haven) |> 
  ggplot(aes(x = initiation_year, y = n, color = any_tax_haven)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c("FALSE" = "grey40", "TRUE" = "red"),
    labels = c("No tax haven", "Tax haven involved")
  ) +
  labs(
    x = "Initiation year",
    y = "Number of deals",
    color = NULL,
    title = "Temporal distribution of land deals by tax haven involvement"
  ) +
  theme_minimal()
```

### Top 10 Tax Haven Investor Countries

#### Which tax haven countries exert the strongest influence on the global land investment network?

```{r}
#| label: fig-Top10-taxhaven-investor-countries
#| fig-cap: "Top 10 Tax Haven Investor Countries(appearances)"

desag |>
  filter(is_tax_haven == TRUE) |>
  count(country_iso2, name = "n_investors") |>
  left_join(tax_haven |> select(iso2, country_name), 
            by = c("country_iso2" = "iso2")) |>
  mutate(country_name = if_else(is.na(country_name), country_iso2, country_name)) |>
  mutate(country_name = forcats::fct_reorder(country_name, n_investors)) |>
  top_n(10, n_investors) |>
  ggplot(aes(x = country_name, y = n_investors)) +
  geom_col(fill = "red") +
  coord_flip() +
  labs(
    x = "Investor Country",
    y = "Number of tax haven investors",
    title = "Top 10 Tax Haven Investor Countries"
  ) +
  theme_minimal()
```

#### Which countries provide the largest number of offshore entities?

```{r}
#| label: fig-Top10-taxhaven-investor-countries-unique
#| fig-cap: "Top 10 Tax Haven Investor Countries(Unique Investors)"

desag |>
  filter(is_tax_haven == TRUE) |>
  distinct(bvd_id_number, country_iso2) |>   # 🔥 unique investor × country
  count(country_iso2, name = "n_unique_investors") |> 
  left_join(tax_haven |> select(iso2, country_name),
            by = c("country_iso2" = "iso2")) |> 
  mutate(
    country_name = if_else(is.na(country_name), country_iso2, country_name),
    country_name = forcats::fct_reorder(country_name, n_unique_investors)
  ) |> 
  top_n(10, n_unique_investors) |> 
  ggplot(aes(x = country_name, y = n_unique_investors)) +
  geom_col(fill = "red") +
  coord_flip() +
  labs(
    x = "Investor Country",
    y = "Number of unique tax haven investors",
    title = "Top 10 Tax Haven Investor Countries (Unique Investors)"
  ) +
  theme_minimal()

```

-   **Singapore is dominant in both frequency and breadth**, meaning it is both a *major offshore incorporation hub* **and** home to *highly active entities repeatedly participating in deals*.

-   **Some jurisdictions (e.g., British Virgin Islands, Cyprus)** show **high appearance counts relative to the number of unique investors**, suggesting **a small number of extremely active offshore structures**.

-   **Other jurisdictions (e.g., Netherlands, Mauritius)** host **many distinct offshore investors**, but each investor tends to participate in fewer deals.\
    → This pattern reflects **distributed incorporation** rather than intensive participation.

-   The comparison reveals **two different modes of offshore involvement** in land investments: **Concentrated influence** (few investors, highly active — e.g., Cyprus, BVI); **Broad-based incorporation** (many investors, moderate activity — e.g., the Netherlands, Mauritius).

### Top 10 target countries for tax haven investors

#### Which countries attract the most investment from tax haven investors?

```{r}
#| label: fig-Top10-target-for-taxhaven
#| fig-cap: "Top 10 Target Countries for Tax Haven Investors"

desag |>
  filter(is_tax_haven == TRUE) |>
  count(target_country_code_alpha2, name = "n_investments") |>
  left_join(
    countries_sf_all |> select(ISO, COUNTRY),
    by = c("target_country_code_alpha2" = "ISO")
  ) |>
  mutate(COUNTRY = if_else(is.na(COUNTRY),
                                target_country_code_alpha2, COUNTRY)) |>
  mutate(COUNTRY = forcats::fct_reorder(COUNTRY, n_investments)) |>
  top_n(10, n_investments) |>
  ggplot(aes(x = COUNTRY, y = n_investments)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    x = "Target Country",
    y = "Number of investments involving tax haven investors",
    title = "Top 10 Target Countries for Tax Haven Investors"
  ) +
  theme_minimal()
```

#### Which countries' land transactions are most affected by tax havens?

```{r}
#| label: fig-Top10-target-for-taxhaven-mostaffected
#| fig-cap: "Top 10 Target Countries for Tax Haven Investors (Deal-level)"

desag |>
  filter(is_tax_haven == TRUE) |>
  distinct(deal_id, target_country_code_alpha2) |>  # 🔥 deal level
  count(target_country_code_alpha2, name = "n_deals") |>
  left_join(
    countries_sf_all |> select(ISO, COUNTRY),
    by = c("target_country_code_alpha2" = "ISO")
  ) |>
  mutate(COUNTRY = if_else(is.na(COUNTRY),
                           target_country_code_alpha2, COUNTRY)) |>
  mutate(COUNTRY = forcats::fct_reorder(COUNTRY, n_deals)) |>
  top_n(10, n_deals) |>
  ggplot(aes(x = COUNTRY, y = n_deals)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    x = "Target Country",
    y = "Number of deals involving tax haven investors",
    title = "Top 10 Target Countries for Tax Haven Investors (Deal-level)"
  ) +
  theme_minimal()

```

### Tax haven investments by sector

```{r}
#| label: cat_current_intention_of_investment
#| 
#=========================================================
# ============ classify_intention ========================
# 1) Preprocessing------------------------------------------------------------
# Standardize case, remove spaces, and correct the misspelling CONVERSATION to CONSERVATION.
desag <- desag |>
  mutate(
    current_intention_of_investment = current_intention_of_investment |>
      toupper() |> str_trim(),
    current_intention_of_investment = str_replace_all(
      current_intention_of_investment,
      "\\bCONVERSATION\\b", "CONSERVATION"
    )
  )

# 2) Definition of classification detection function-----------------------------
# whether it contains certain keywords
has_any <- function(x, patterns) 
  str_detect(x, str_c("\\b(", str_c(patterns, collapse="|"), ")\\b"))

# 3) Rule: Process by priority using CASE WHEN (stop upon first match)-----------
desag <- desag |>
  mutate(
    intention_investment_cat = case_when(
      is.na(current_intention_of_investment) | current_intention_of_investment == "" ~ "NA",
      
      has_any(current_intention_of_investment,
              c("BIOFUELS","RENEWABLE_ENERGY","WIND_FARM","SOLAR_PARK",
                "OIL_GAS_EXTRACTION","BIOMASS_ENERGY_GENERATION")) ~ "Energy",
      
      has_any(current_intention_of_investment, c("MINING")) ~ "Mining",
      
      has_any(current_intention_of_investment,
              c("FOOD_CROPS","NON_FOOD_AGRICULTURE","LIVESTOCK","FODDER",
                "AGRICULTURE_UNSPECIFIED")) ~ "Agriculture",
      
      has_any(current_intention_of_investment,
              c("CONSERVATION","CARBON")) ~ "Conservation and Carbon",
      
      has_any(current_intention_of_investment,
              c("FOREST_LOGGING","TIMBER_PLANTATION","FORESTRY_UNSPECIFIED")) ~ "Forestry",
      
      has_any(current_intention_of_investment, c("TOURISM")) ~ "Tourism",
      
      has_any(current_intention_of_investment, c("INDUSTRY")) ~ "Industry",
      
      has_any(current_intention_of_investment, c("LAND_SPECULATION")) ~ "Land speculation",
      
      has_any(current_intention_of_investment, c("OTHER")) ~ "Other",
      
      TRUE ~ "Other" 
    )
  )
```

```{r}
#| label: fig-taxhaven-by-sector
#| fig-cap: "Tax Haven Investments by Sector (Share)"

value_summary(desag, "intention_investment_cat")

# Compute sector share of tax haven involvement
sector_taxhaven <- desag |>
  group_by(intention_investment_cat) |>
  summarise(
    total = n(),
    taxhaven = sum(is_tax_haven, na.rm = TRUE),
    share = taxhaven / total,
    .groups = "drop"
  ) |>
  arrange(desc(share))

# Plot percentage involvement
sector_taxhaven |>
  ggplot(aes(
    x = forcats::fct_reorder(intention_investment_cat, share),
    y = share
  )) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Investment sector",
    y = "Share of tax-haven investors",
    title = "Tax Haven Investments by Sector (Share)"
  ) +
  theme_minimal()


```

```{r}
sector_stacked <- desag |>
  group_by(intention_investment_cat, is_tax_haven) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(intention_investment_cat) |>
  mutate(share = n / sum(n))

sector_stacked |>
  ggplot(aes(
    x = intention_investment_cat,
    y = share,
    fill = is_tax_haven
  )) +
  geom_col(position = "fill") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(
    values = c("FALSE" = "grey70", "TRUE" = "red"),
    labels = c("No tax haven", "Tax haven involved")
  ) +
  labs(
    x = "Investment sector",
    y = "Share of investor appearances",
    fill = NULL,
    title = "Tax Haven Involvement by Sector (Proportional Stacked Bar)"
  ) +
  theme_minimal()

```

```{r}
# Step 1: Compute counts by (tax haven investor country × sector)
country_sector <- desag |>
  filter(is_tax_haven == TRUE) |>
  count(country_iso2, intention_investment_cat, name = "n") 

# Step 2: Compute total per country
country_total <- country_sector |>
  group_by(country_iso2) |>
  summarise(total = sum(n), .groups = "drop")

# Step 3: Select top 10 tax haven investor countries
top10_countries <- country_total |>
  top_n(10, total) |>
  arrange(desc(total)) |>
  pull(country_iso2)

# Step 4: Filter country_sector to only these 10 countries
plot_df <- country_sector |>
  filter(country_iso2 %in% top10_countries) |>
  left_join(country_total, by = "country_iso2") |>
  # reorder for nice plotting
  mutate(country_iso2 = forcats::fct_reorder(country_iso2, total))

# Step 5: Add readable country names
plot_df <- plot_df |>
  left_join(tax_haven |> select(iso2, country_name),
            by = c("country_iso2" = "iso2")) |>
  mutate(country_name = if_else(is.na(country_name),
                                country_iso2,
                                country_name)) |>
  mutate(country_name = forcats::fct_reorder(country_name, total))
  

# Step 6: Plot stacked bar chart

ggplot(plot_df, aes(
  x = country_name,
  y = n,
  fill = intention_investment_cat
)) +
  geom_col() +
  coord_flip() +
  labs(
    x = "Tax Haven Investor Country",
    y = "Number of investor appearances",
    fill = "Sector",
    title = "Sector Composition of Top 10 Tax Haven Investor Countries",
    subtitle = "Stacked bar: colour indicates investment sector"
  ) +
  theme_minimal() +
  theme(legend.position = "right")

```

### map

## Network Analysis

-   Network graph

    -   highlighting tax haven nodes

-   Table of top 10 countries by centrality (degree, betweenness)

    -   evaluate the bridging role of tax havens;

1.  network map: Country-to-country investment network -\> Illustrating the global investment structure and the locations of tax haven nodes

2.  table: Top 10 countries by betweenness centrality -\> Identifying which tax havens serve as key bridge nodes is crucial.

3.  network map: Regional clusters of tax havens -\> Demonstrating the clustering characteristics of Caribbean, European and Asian tax havens

4.  line: Network density evolution of tax haven participation -\> Monitor network connectivity trends over time

## Role-Level Analysis

-   Examine how tax haven entities appear across **investment roles**:

    -   Operating Company (OC)

    -   Shareholder (SH)

    -   Controlling Shareholder (CSH)

    -   Global Ultimate Owner (GUO)

-   Summary table: Number of tax haven entities by role. -\> Concentration of tax havens across OC, SH, CSH, and GUO tiers

-   Sankey diagram: Country → Role → Target country. -\> Illustrating investment pathways and the role of tax havens within them

## Impact Assessment
