# Analysis

```{r init}
#| label: Analysis

source("globals.R")
```

## Data Overview

### Data availability

```{r}
#| label: check-availability

# check data.csv 
data_availability(
  data_name = "data",
  fields = c("initiation_year", "current_intention_of_investment", "investment_type","world_region","country_iso3_code","country_iso_code")
)

# check desag.csv
data_availability(
  data_name = "desag",
  fields = c("role", "country_iso3", "target_country_iso3", "is_tax_haven", "current_implementation_status", "initiation_year")
)

```

```{r}
#| label: value-summary

range(data$initiation_year, na.rm = TRUE)
value_summary(data, "current_implementation_status")
value_summary(data, "investment_type")

value_summary(desag, "is_tax_haven")
value_summary(desag, "role")

value_summary(investors, "from_tax_haven")
```

## Tax Havens in the Investment Network

### How often do tax havens appear

```{r}
#| label: create-deals-level-involve-TaxHaven

deal_taxhaven <- desag |>
  group_by(deal_id) |>
  summarise(
    num_tax = sum(is_tax_haven, na.rm = TRUE),
    any_tax_haven = num_tax > 0,
    current_implementation_status = first(current_implementation_status),
    initiation_year = first(initiation_year),
    .groups = "drop"
  )

value_summary(deal_taxhaven, "any_tax_haven")
value_summary(deal_taxhaven, "num_tax")

```

```{r}
#| label: create-investor-level-involve-TaxHaven

investor_taxhaven <- desag |>
  group_by(bvd_id_number) |>
  summarise(
    investor_country = paste(unique(country_iso3), collapse = ", "),
    roles = paste(unique(role), collapse = ", "),
    num_deals = n_distinct(deal_id),
    is_tax_haven_investor = any(is_tax_haven, na.rm = TRUE),
    .groups = "drop"
  )

value_summary(investor_taxhaven, "is_tax_haven_investor")
```

#### % of deals involving at least one tax-haven investor

```{r}
#| label: deals-involving-taxhaven-investor

deal_taxhaven |>
  summarise(
    total_deals = n(),
    tax_haven_deals    = sum(any_tax_haven),
    share       = scales::percent(tax_haven_deals / total_deals, accuracy = 0.1)
  )|>
  knitr::kable(
    caption = "Deals involving at least one tax-haven investor",
    align = "c"
  ) |>
  kableExtra::kable_styling(full_width = FALSE, position = "center")
```

-   Share of land deals currently ***in operation*** that involve tax-haven investors

```{r}
#| label: deals-involving-taxhaven-investor-in-operation

deal_taxhaven |>
  filter(current_implementation_status == "In operation") |>
  summarise(
    total_in_operation = n(),
    tax_haven_deals           = sum(any_tax_haven),
    share              = scales::percent(tax_haven_deals / total_in_operation, accuracy = 0.1)
  ) |>
  knitr::kable(
    caption = "Deals involving at least one tax-haven investor (In operation)",
    align = "c"
  ) |>
  kableExtra::kable_styling(full_width = FALSE, position = "center")

```

To quantify the involvement of tax-haven–registered investors, we aggregated the investor-level information in *desag_data* to the deal level. **A deal is considered to involve a tax haven if at least one investor within its ownership chain is registered in a jurisdiction listed as a tax haven.**

The results show that **1318** deals (representing **24.7%** of all deals) involve at least one tax-haven investor. This indicates that tax-haven jurisdictions play a non-negligible role in land investment structures, either through shareholders, controlling shareholders, or ultimate owners.

When considering only deals currently ***in operation***, **946** deals out of **3446** are associated with tax-haven investors, corresponding to **27.5%**.

#### **Tax haven land deal time trend**

```{r}
#| label: fig-deals-taxhaven-vs-total
#| fig-cap: "Evolution of total deals and tax-haven-involved deals over time"

# ----- 1. Prepare the data -----
deal_year_summary <- deal_taxhaven |>
  group_by(initiation_year) |>
  summarise(
    tax_haven_deals = sum(any_tax_haven, na.rm = TRUE),
    total_deals = n(),
    .groups = "drop"
  )

# Convert to long format for ggplot
deal_plot_data <- deal_year_summary |>
  tidyr::pivot_longer(
    cols = c(tax_haven_deals, total_deals),
    names_to = "type",
    values_to = "n"
  ) |>
  mutate(
    type = recode(type,
                  "tax_haven_deals" = "Tax-haven deals",
                  "total_deals"     = "Total deals")
  )

# ----- 2. Find earliest tax-haven deal year -----
first_tax_year <- deal_year_summary |>
  filter(tax_haven_deals > 0) |>
  slice_min(initiation_year) |>
  pull(initiation_year)

# Value at that point
first_tax_value <- deal_year_summary |>
  filter(initiation_year == first_tax_year) |>
  pull(tax_haven_deals)

# ----- 3. Plot -----
ggplot(deal_plot_data,
       aes(x = initiation_year,
           y = n,
           color = type)) +
  
  geom_line(linewidth = 1) +
  
  # Colors
  scale_color_manual(
    values = c(
      "Tax-haven deals" = "#E97437",
      "Total deals"     = "black"
    )
  ) +
  
  # Label the first tax-haven deal
  geom_point(
    data = tibble(initiation_year = first_tax_year,
                  n = first_tax_value),
    aes(x = initiation_year, y = n),
    color = "#E97437",
    size = 3
  ) +
  
  geom_text(
    data = tibble(initiation_year = first_tax_year,
                  n = first_tax_value),
    aes(label = first_tax_year),
    color = "#E97437",
    vjust = -1,
    fontface = "bold",
    size = 5
  ) +
  
  labs(
    title = "Deals: Total vs Tax-haven Involvement",
    x = "Initiation year",
    y = "Number of deals",
    color = NULL
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 18)
  )

# ----- ----- -----
# ----- ----- -----
# Filter data from 1990 onward
deal_plot_data_filtered <- deal_plot_data %>%
  filter(initiation_year >= 1990)

ggplot(deal_plot_data_filtered,
       aes(x = initiation_year,
           y = n,
           color = type)) +
  
  geom_line(linewidth = 1) +
  
  scale_color_manual(
    values = c(
      "Tax-haven deals" = "#E97437",
      "Total deals"     = "black"
    )
  ) +
  
  labs(
    title = "Deals: Total vs Tax-haven Involvement",
    x = "Initiation year (≥ 1990)",
    y = "Number of deals",
    color = NULL
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 18)
  )


```

#### % of investors registered in TH

```{r}
#| label: fig-unique-investor-taxhaven-share
#| fig-cap: "Number and share of unique investors registered in tax havens"

investor_summary <- investor_taxhaven |>
  count(is_tax_haven_investor) |>
  mutate(
    share = n / sum(n),
    label = if_else(is_tax_haven_investor, 
                    "Tax Haven Investor", 
                    "Non-Tax Haven Investor"),
    num_investors = paste0(n, "\n", scales::percent(share, accuracy = 0.1))
  )

investor_summary |>
  ggplot(aes(x = 1.8, y = n, fill = label)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 0) +
  scale_fill_manual(values = c("Tax Haven Investor" = "#E97437",
                               "Non-Tax Haven Investor" = "#3F7B97")) +
  geom_text(aes(label = num_investors),
            position = position_stack(vjust = 0.5),
            size = 4, color = "white", lineheight = 0.9) +
  theme_void() +
  xlim(0.5, 2.5) +
  labs(fill = NULL)
```

#### Average Number of Deals per Investor

```{r}
#| label: Average-Number-of-Deals-per-Investor
#| fig-cap: "TH vs Non-TH Investors: Activity Level"

investor_taxhaven |>
  group_by(is_tax_haven_investor) |>
  summarise(
    avg_deals = mean(num_deals),
    median_deals = median(num_deals),
    max_deals = max(num_deals),
    total_deal_appearances = sum(num_deals),
    n_investors = n(),
    .groups = "drop"
  )

# Overall distribution 
value_summary(investor_taxhaven,"num_deals" )

value_summary(investor_taxhaven |> filter(is_tax_haven_investor == TRUE), "num_deals")

value_summary(investor_taxhaven |> filter(is_tax_haven_investor == FALSE), "num_deals")

```

### ！！！地理方面的考虑一下是做地图还是按照这个Geographic Origins of Tax Haven Investors

#### Top 10 Tax Haven Investor Countries

-   各离岸司法辖区出现频次

-   区分两类模式： ① 高度活跃型避税天堂（如 BVI、塞浦路斯：少量实体但参与交易频繁） ② 广泛设立型避税天堂（如荷兰、新加坡：投资者数量多但每个涉及交易有限）

-   解释不同模式对土地投资网络的影响

```{r}
#| label: create-Top10-taxhaven-investor-countries

th_appearances <- desag |>
  filter(is_tax_haven) |>
  count(country_iso3, name = "n_appearances")

th_unique <- desag |>
  filter(is_tax_haven) |>
  distinct(bvd_id_number, country_iso3) |>
  count(country_iso3, name = "n_unique")

th_profile <- th_appearances |>
  left_join(th_unique, by = "country_iso3") |>
  mutate(
    avg_activity = n_appearances / n_unique  # Average number of transactions per investor
  ) |>
  arrange(desc(avg_activity))
```

```{r}
#| label: fig-Top10-taxhaven-investor-countries
#| fig-cap: "Top 10 Tax Haven Investor Countries"

p1 <- th_appearances |>
  left_join(tax_haven |> select(iso3, country_name),
            by = c("country_iso3" = "iso3")) |>
  mutate(country_name = if_else(is.na(country_name), country_iso3, country_name),
         country_name = forcats::fct_reorder(country_name, n_appearances)) |>
  top_n(10, n_appearances) |>
  ggplot(aes(x = country_name, y = n_appearances)) +
  geom_col(fill = "red") +
  coord_flip() +
  labs(
    x = "Investor Country",
    y = "Number of investor appearances",
    title = "Top 10 active tax haven (frequency of occurrence)"
  ) +
  theme_minimal()

p2 <- th_unique |>
  left_join(tax_haven |> select(iso3, country_name),
            by = c("country_iso3" = "iso3")) |>
  mutate(country_name = if_else(is.na(country_name), country_iso3, country_name),
         country_name = forcats::fct_reorder(country_name, n_unique)) |>
  top_n(10, n_unique) |>
  ggplot(aes(x = country_name, y = n_unique)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    x = "Investor Country",
    y = "Number of unique investors",
    title = "Top 10 active tax haven (Number of Unique Investors)"
  ) +
  theme_minimal()

p1 + p2 + plot_layout(ncol = 1)
```

```{r}
#| label: Identifying-'Highly Active'-vs-'Broadly Established'-TaxHavens

th_profile |>
  mutate(
    category = case_when(
      avg_activity >= 3 ~ "Highly Active",
      TRUE ~ "Broadly Established"
    )
  )
```

-   **Singapore is dominant in both frequency and breadth**, meaning it is both a *major offshore incorporation hub* **and** home to *highly active entities repeatedly participating in deals*.

-   **Some jurisdictions (e.g., British Virgin Islands, Cyprus)** show **high appearance counts relative to the number of unique investors**, suggesting **a small number of extremely active offshore structures**.

-   **Other jurisdictions (e.g., Netherlands, Mauritius)** host **many distinct offshore investors**, but each investor tends to participate in fewer deals.\
    → This pattern reflects **distributed incorporation** rather than intensive participation.

-   The comparison reveals **two different modes of offshore involvement** in land investments: **Concentrated influence** (few investors, highly active — e.g., Cyprus, BVI); **Broad-based incorporation** (many investors, moderate activity — e.g., the Netherlands, Mauritius).

## Position of Tax Haven Entities in Ownership Chains

### Role distribution: Where do they appear in the chain?

```{r}
#| label: role-by-taxhaven-appearance
#| fig-cap: "Number of appearances and proportion of tax haven investors by investment role"

role_appearance <- desag |>
  filter(!is.na(role)) |>
  mutate(
    is_tax_haven = if_else(is_tax_haven, "Tax haven investors", "Non-tax haven investors")
  ) |>
  count(role, is_tax_haven, name = "n") |>
  group_by(role) |>
  mutate(
    total_role = sum(n),
    share = n / total_role
  ) |>
  ungroup()  

# result table
role_appearance |>
  arrange(role, desc(is_tax_haven)) |>
  knitr::kable(
    caption = "Occurrence and Proportion of Tax Haven and Non-Tax Haven Investors by Investment Role"
  )

# Visualisation: Proportion of tax havens versus non-tax havens under each role
role_appearance |>
  ggplot(aes(x = role, y = share, fill = is_tax_haven)) +
  geom_col(position = "fill") +
  
  # Add percentage text 
  geom_text(aes(label = scales::percent(share, accuracy = 1)),
            position = position_fill(vjust = 0.5),
            color = "white",
            size = 3.5) +
  
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  
  # Modify the horizontal axis labels
  scale_x_discrete(labels = c(
    "Operating Company" = "OC",
    "Shareholder" = "SH"
  )) +
  
  scale_fill_manual(
    values = c(
      "Tax haven investors"  = "#E97437",  # Tax-haven investors
      "Non-tax haven investors" = "#3F7B97"   # Non-tax haven investors
    ),
    labels = c(
      "Tax haven investors"  = "Tax haven investors",
      "Non-tax haven investors" = "Non-tax haven investors"
    )
  ) +
  
  labs(
    x = "Investment role",
    y = "Percentage of times the character appears",
    fill = NULL,
    title = "Relative Proportion of Tax Haven Investors Across Different Ownership Roles"
  ) +
  theme_minimal()

```

#### vertical distance of TH from OC

structural level indicator, to confirm the question: Do tax havens primarily appear at the upper levels of the ownership chain (CSH, GUO) rather than the lower levels (OC, SH)?

```{r}

role_levels <- c(
  "Operating Company"  = 0,
  "Shareholder"  = 1,
  "ISH" = 1,
  "DUO" = 2,
  "CSH" = 3,
  "GUO" = 4
)

desag <- desag |>
  mutate(
    role_level = role_levels[role]  
  )

vertical_summary <- desag |>
  group_by(is_tax_haven) |>
  summarise(
    avg_level = mean(role_level, na.rm = TRUE),
    median_level = median(role_level, na.rm = TRUE),
    sd_level = sd(role_level, na.rm = TRUE),
    n = n()
  )
vertical_summary


ggplot(vertical_summary,
       aes(x = is_tax_haven, y = avg_level, fill = is_tax_haven)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = round(avg_level, 2)), vjust = -0.3, size = 5) +
  scale_fill_manual(values = c("#3F7B97", "#E97437")) +
  labs(
    title = "Average Ownership-Chain Level: Tax Haven vs Other Investors",
    x = "Is Tax Haven Deal",
    y = "Average vertical level"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")

```

## Role-Specific Patterns by Tax Haven Clusters

### **A: regional cluster**

**Step 1: Define regional Cluster**

```{r}
tax_haven_clusters <- tibble::tribble(
  ~country_name,                   ~iso3,  ~cluster,

  # ------------------- Africa -------------------
  "Mauritius",                    "MUS",   "Africa",

  # ------------------- Caribbean ----------------
  "Bahamas",                      "BHS",   "Caribbean",
  "Bermuda",                      "BMU",   "Caribbean",
  "British Virgin Islands",       "VGB",   "Caribbean",
  "Cayman Islands (the)",         "CYM",   "Caribbean",
  "Panama",                       "PAN",   "Caribbean",

  # ------------------- Eastern Asia --------------
  "Hong Kong",                    "HKG",   "Eastern Asia",
  "Singapore",                    "SGP",   "Eastern Asia",

  # ------------------- Europe --------------------
  "Belgium",                      "BEL",   "Europe",
  "Cyprus",                       "CYP",   "Europe",
  "Estonia",                      "EST",   "Europe",
  "Hungary",                      "HUN",   "Europe",
  "Ireland",                      "IRL",   "Europe",
  "Liechtenstein",                "LIE",   "Europe",
  "Luxembourg",                   "LUX",   "Europe",
  "Malta",                        "MLT",   "Europe",
  "Netherlands (Kingdom of the)", "NLD",   "Europe",
  "Switzerland",                  "CHE",   "Europe",

  # ------------------- Western Asia --------------
  "Lebanon",                      "LBN",   "Western Asia",
  "United Arab Emirates (the)",   "ARE",   "Western Asia"
)

```

```{r}
th_with_cluster <- desag |>
  left_join(tax_haven_clusters, by = c("country_iso3" = "iso3"))|>
  filter(!is.na(cluster))

role_distribution <- th_with_cluster |>
  count(cluster, role) |>
  group_by(cluster)  |>
  mutate(pct = n / sum(n)) |>
  ungroup()

```

```{r}

ggplot(role_distribution, aes(x = cluster, y = pct, fill = role)) +
  geom_col(width = 0.7, alpha = 0.85) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(
    values = c(
      "Operating Company"  = "#A6CEE3",
      "Shareholder"  = "#1F78B4",
      "ISH" = "#B2DF8A",
      "CSH" = "#33A02C",
      "DUO" = "#FB9A99",
      "GUO" = "#E31A1C"
    )
  ) +
  labs(
    title = "Roles played by Tax Havens in the ownership chain",
    subtitle = "Different tax haven clusters occupy distinct structural positions",
    x = NULL,
    y = "Share (%)",
    fill = "Role"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 10),
    plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40"),
    axis.text.x = element_text(size = 10, face = "bold", margin = margin(t = 5)),
    axis.text.y = element_text(size = 10)
  )

```

```{r}
ggplot(role_distribution, aes(x = cluster, y = n, fill = role)) +
  geom_col(width = 0.7, alpha = 0.85) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(
    values = c(
      "Operating Company"  = "#A6CEE3",
      "Shareholder"  = "#1F78B4",
      "ISH" = "#B2DF8A",
      "CSH" = "#33A02C",
      "DUO" = "#FB9A99",
      "GUO" = "#E31A1C"
    )
  ) +
  labs(
    title = "Number of Investor Appearances by Role within Tax Haven Clusters",
    subtitle = "Different tax haven clusters occupy distinct structural positions",
    x = NULL,
    y = "Number of investor occurrences",
    fill = "Role"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 10),
    plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 10, color = "grey40"),
    axis.text.x = element_text(size = 10, face = "bold", margin = margin(t = 5)),
    axis.text.y = element_text(size = 10)
  )

```

### B: Do clusters truly exist? (Exploratory analysis)

```{r}
# 只保留避税天堂投资者
th_desag <- desag |>
  filter(is_tax_haven == TRUE)

# 检查一下有哪些避税天堂国家和角色
th_desag |>
  count(country_iso3, role) |>
  arrange(desc(n)) |>
  print(n = 50)
```

## 步骤 1：为每个避税天堂国家做一个“角色画像”

我们要给每个 TH 国家算出：

-    作为 OC 的次数 / 总次数

-    作为 SH（包括 ISH） 的次数 / 总次数

-   作为 CSH 的次数 / 总次数

-   作为 GUO 的次数 / 总次数

```{r}
th_role_profile <- th_desag |>
  group_by(country_iso3, role) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(country_iso3) |>
  mutate(
    total_n = sum(n),
    share = n / total_n
  ) |>
  ungroup()

th_role_profile
```

```{r}
th_matrix <- th_role_profile |>
  select(country_iso3, role, share) |>
  pivot_wider(
    names_from = role,
    values_from = share,
    values_fill = 0
  )

th_matrix

```

## 步骤 2：过滤掉样本太少的国家（避免噪音）

如果某个 TH 国家在你的数据里只出现了 3 次，它的比例会很不稳定。\
所以通常会加一个阈值，例如：只保留 total_n ≥ 30 的国家。

```{r}
country_size <- th_desag |>
  count(country_iso3, name = "total_n")

th_matrix_filtered <- th_matrix |>
  left_join(country_size, by = "country_iso3") |>
  filter(total_n >= 50) |>
  select(-total_n)

th_matrix_filtered

```

## 步骤 3：对国家做聚类（层次聚类）

### 3.1 建距离矩阵 + 聚类

```{r}
# 去掉 country 列，留下数值矩阵
mat <- th_matrix_filtered |>
  select(-country_iso3) |>
  as.matrix()

# 可选：标准化（这里 share 本身已经是比例，一般可以不 scale）
# mat_scaled <- scale(mat)

# 计算距离（欧式）
dist_mat <- dist(mat)

# 层次聚类
hc <- hclust(dist_mat, method = "ward.D2")

# 画树状图
plot(hc, labels = th_matrix_filtered$country_iso3, main = "TH countries clustered by role profiles")

```

### 3.2 选择一个合理的 cluster 数目

你可以先直观地“看”，比如先试试切成 3 类、4 类：

```{r}
# 比如先切成 3 类
k <- 3
cluster_assign <- cutree(hc, k = k)

th_clusters <- th_matrix_filtered |>
  mutate(cluster_id = factor(cluster_assign))

th_clusters

th_clusters |>
  arrange(cluster_id) |>
  select(country_iso3, cluster_id)

```

## 步骤 4：用可视化辅助理解 cluster

### 4.1 在角色空间里看 cluster

```{r}

th_role_profile_clustered <- th_role_profile |>
  inner_join(th_clusters |> select(country_iso3, cluster_id), by = "country_iso3")

# 画 heatmap：各 cluster 的角色结构
th_role_profile_clustered |>
  group_by(cluster_id, role) |>
  summarise(avg_share = mean(share), .groups = "drop") |>
  ggplot(aes(x = role, y = cluster_id, fill = avg_share)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(
    x = "Role",
    y = "Cluster",
    fill = "Average share",
    title = "Average role share per cluster (data-driven clusters)"
  ) +
  theme_minimal()

```

-   欧洲集群（荷兰、卢森堡、塞浦路斯）倾向于出现在高层控股结构中

-   加勒比集群（BVI、开曼等）常处于 Ultimate Owner 或多层控股链条中

-   亚洲集群（新加坡、香港）多作为区域性金融节点

-   解释不同集群在链条中的功能差异

## Sectoral and Value Chain Exposure to Tax Havens不同行业（价值链）对避税天堂的暴露程度

### Sector-Level Exposure

Which sectors are most likely to involve tax havens?

```{r}
# tax haven deal's deal_id
desag_with_cluster <- desag |>
  left_join(tax_haven_clusters, by = c("country_iso3" = "iso3"))

deals_with_TH <- desag_with_cluster |>
  filter(!is.na(cluster)) |>
  distinct(deal_id)|>
  pull(deal_id)

sector_total <- desag |>
  distinct(deal_id, intention_investment_cat)


sector_TH <- sector_total |>
  mutate(tax_haven_involved = deal_id %in% deals_with_TH) |>
  group_by(intention_investment_cat) |>
  summarise(
    total_deals = n(),
    th_deals = sum(tax_haven_involved),
    pct_th = th_deals / total_deals
  ) |>
  arrange(desc(pct_th))


ggplot(sector_TH, aes(x = reorder(intention_investment_cat, pct_th), 
                      y = pct_th)) +
  geom_col(fill = "#E97437", alpha = 0.8) +
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip() +
  labs(
    title = "Which sectors are most associated with tax havens?",
    subtitle = "Share of deals involving at least one tax haven investor",
    x = "Sector (intention_investment_cat)",
    y = "Share of tax-haven-involved deals (%)"
  ) +
  theme_minimal(base_size = 14)

```

### Sector × Ownership Position

What role do tax havens play in different value chains?

```{r}
desag_TH <- desag_with_cluster |>
  filter(!is.na(cluster), !is.na(role), !is.na(intention_investment_cat))

sector_role <- desag_TH |>
  distinct(deal_id, intention_investment_cat, role) |>
  count(intention_investment_cat, role) |>
  group_by(intention_investment_cat) |>
  mutate(pct = n / sum(n)) |>
  ungroup()

```

```{r}

role_order <- c("Operating Company", "Shareholder", "ISH", "CSH", "DUO", "GUO")

sector_role$role <- factor(sector_role$role, levels = role_order)

ggplot(sector_role, aes(x = intention_investment_cat, y = pct, fill = role)) +
  geom_col(alpha = 0.9, width = 0.75) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(
    values = c(
      "Operating Company" = "#A6CEE3",
      "Shareholder"       = "#1F78B4",
      "ISH"               = "#B2DF8A",
      "CSH"               = "#33A02C",
      "DUO"               = "#FB9A99",
      "GUO"               = "#E31A1C"
    )
  ) +
  labs(
    title = "Roles of Tax Havens across Value Chains",
    subtitle = "Role composition of tax-haven investors within each sector",
    x = "Sector (intention_investment_cat)",
    y = "Share of roles (%)",
    fill = "Role"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 25, hjust = 1),
    legend.position = "bottom"
  )

```

### Sector × Tax Haven Clusters

Which value chains are most strongly associated with which clusters?

```{r}
# Obtain the deals each cluster participates in (avoiding duplicate counting)
cluster_deals <- desag_with_cluster |>
  filter(!is.na(cluster)) |>           
  distinct(cluster, deal_id)            

deal_sector <- desag |>
  distinct(deal_id, intention_investment_cat)

cluster_sector <- cluster_deals |>
  left_join(deal_sector, by = "deal_id")

cluster_sector_count <- cluster_sector |>
  group_by(cluster, intention_investment_cat) |>
  summarise(
    th_deals = n(),         # cluster 在该 sector 参与的 deal 数
    .groups = "drop"
  )

sector_total <- desag |>
  distinct(deal_id, intention_investment_cat) |>
  count(intention_investment_cat, name = "total_deals")

cluster_sector_stats <- cluster_sector_count |>
  left_join(sector_total, by = "intention_investment_cat") |>
  mutate(pct_th = th_deals / total_deals) |>     # 占比
  arrange(desc(pct_th))

```

```{r}


ggplot(cluster_sector_stats,
       aes(x = intention_investment_cat,
           y = cluster,
           fill = pct_th)) +
  geom_tile(color = "white", linewidth = 0.8) +
  scale_fill_viridis_c(option = "magma", direction = -1,
                       labels = scales::percent_format()) +
  labs(
    title = "Which value chains are most associated with each Tax Haven Cluster?",
    subtitle = "Share of deals involving each tax haven cluster within each sector",
    x = "Sector (intention_investment_cat)",
    y = "Tax Haven Cluster",
    fill = "Share (%)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    panel.grid = element_blank()
  )

```

### Sector× cluster × role

Do different tax haven clusters play distinct roles across different value chains?

```{r}

# tax haven only
th_data <- desag_with_cluster |>
  filter(!is.na(cluster), !is.na(role), !is.na(intention_investment_cat)) |>
  distinct(cluster, intention_investment_cat, role, deal_id)

sector_role_cluster <- th_data |>
  count(cluster, intention_investment_cat, role) |>
  group_by(cluster, intention_investment_cat) |>
  mutate(pct = n / sum(n)) |>
  ungroup()

```

```{r}

role_order <- c("Operating Company", "Shareholder", "ISH", "CSH", "DUO", "GUO")
sector_role_cluster$role <- factor(sector_role_cluster$role, levels = role_order)

ggplot(sector_role_cluster,
       aes(x = intention_investment_cat, y = pct, fill = role)) +
  geom_col(alpha = 0.9, width = 0.75) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(
    values = c(
      "Operating Company" = "#A6CEE3",
      "Shareholder"       = "#1F78B4",
      "ISH"               = "#B2DF8A",
      "CSH"               = "#33A02C",
      "DUO"               = "#FB9A99",
      "GUO"               = "#E31A1C"
    )
  ) +
  facet_wrap(~ cluster, ncol = 3) +
  labs(
    title = "Roles of Tax Havens Across Value Chains, by Cluster",
    subtitle = "Each cluster shows distinct structural roles within sectors",
    x = "Sector (intention_investment_cat)",
    y = "Role share (%)",
    fill = "Role"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 14),
    plot.title = element_text(face = "bold", size = 14)
  )

```

### Sector × cluster × GUO 

In different value chains, which countries do various tax haven clusters ultimately connect to?

```{r}
th_guo <- desag_with_cluster |>
  filter(!is.na(cluster)) |>               # tax haven investor
  filter(role == "GUO") |>                # 只看最终控制者
  filter(!is.na(intention_investment_cat),
         !is.na(country_name)) |>     # 必要字段
  distinct(cluster, intention_investment_cat, country_name, deal_id)

cluster_sector_guo <- th_guo |> 
  count(cluster, intention_investment_cat, country_name) |> 
  group_by(cluster, intention_investment_cat) |> 
  mutate(pct = n / sum(n)) |>      # 占比（哪个 GUO 来源占主导）
  ungroup()

```

```{r}
library(networkD3)

# 构建节点
nodes <- data.frame(
  name = unique(c(
    cluster_sector_guo$intention_investment_cat,
    cluster_sector_guo$country_name
  ))
)

# 构建连接
links <- cluster_sector_guo %>%
  mutate(
    source = match(intention_investment_cat, nodes$name) - 1,
    target = match(country_name, nodes$name) - 1,
    value = n
  ) %>%
  select(source, target, value)
  
sankeyNetwork(
  Links = links,
  Nodes = nodes,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "name",
  sinksRight = FALSE,
  fontSize = 14,
  nodeWidth = 30
)

```

```{r}

sankey_4d <- desag_with_cluster |>
  filter(!is.na(cluster), !is.na(intention_investment_cat),
         !is.na(role), !is.na(country_name)) |>
  
  # 只计算每个路径在每个 deal 中出现一次
  distinct(cluster, intention_investment_cat, role, country_name, deal_id)|>
  
  count(cluster, intention_investment_cat, role, country_name) |>
  rename(value = n)


nodes <- data.frame(
  name = unique(c(
    sankey_4d$cluster,
    sankey_4d$intention_investment_cat,
    sankey_4d$role,
    sankey_4d$country_name
  ))
)
links1 <- sankey_4d |>
  group_by(cluster, intention_investment_cat) |>
  summarise(value = sum(value), .groups = "drop") |>
  mutate(
    source = match(cluster, nodes$name) - 1,
    target = match(intention_investment_cat, nodes$name) - 1
  ) |>
  select(source, target, value)

links2 <- sankey_4d |>
  group_by(intention_investment_cat, role) |>
  summarise(value = sum(value), .groups = "drop") |>
  mutate(
    source = match(intention_investment_cat, nodes$name) - 1,
    target = match(role, nodes$name) - 1
  ) |>
  select(source, target, value)

links3 <- sankey_4d |>
  group_by(role, country_name) |>
  summarise(value = sum(value), .groups = "drop") |>
  mutate(
    source = match(role, nodes$name) - 1,
    target = match(country_name, nodes$name) - 1
  ) |>
  select(source, target, value)

links_all <- bind_rows(links1, links2, links3)

sankeyNetwork(
  Links = links_all,
  Nodes = nodes,
  Source = "source",
  Target = "target",
  Value = "value",
  NodeID = "name",
  fontSize = 13,
  nodeWidth = 30,
  sinksRight = FALSE
)

```

## 
