# Analysis

```{r init}
#| label: Analysis

source("globals.R")
```

## Data Overview

### **Expected:**

-   fig/text: Data coverage and share of “in operation” deals in the total deals.

-   xxxx unique investors identified, around xxxx% of investors registered in tax havens.

-   fig: Temporal distribution of land deals

### Data availability

```{r}
#| label: check-availability

# check data.csv 
data_availability(
  data_name = "data",
  fields = c("initiation_year", "current_intention_of_investment", "investment_type", "target_country_code_alpha2","world_region")
)

# check desag.csv
data_availability(
  data_name = "desag",
  fields = c("role", "country", "target_country_code_alpha2")
)
```

```{r}
#| label: value-summary

value_summary(data, "current_implementation_status")

value_summary(data, "world_region")
```

### How many deals involve tax havens?

## Distribution Patterns

1.  bar: Share of tax haven vs. non–tax haven investments

2.  line: Evolution of tax haven investments over time -\> Analysis of whether the use of tax havens has increased over time

3.  bar: Top 10 tax haven investor countries -\> Identify the primary source countries of tax havens

4.  bar: Top 10 target countries for tax haven investors -\> Identify the principal countries of investment

5.  Map: Global investment flows from tax havens to target countries -\> Visualising the global investment pathways of tax havens

6.  bar: Tax haven investments by sector -\> Comparing the proportion of tax havens across different investment sectors

## Network Analysis

-   Network graph

    -   highlighting tax haven nodes

-   Table of top 10 countries by centrality (degree, betweenness)

    -   evaluate the bridging role of tax havens;

1.  network map: Country-to-country investment network -\> Illustrating the global investment structure and the locations of tax haven nodes

2.  table: Top 10 countries by betweenness centrality -\> Identifying which tax havens serve as key bridge nodes is crucial.

3.  network map: Regional clusters of tax havens -\> Demonstrating the clustering characteristics of Caribbean, European and Asian tax havens

4.  line: Network density evolution of tax haven participation -\> Monitor network connectivity trends over time

## Role-Level Analysis

-   Examine how tax haven entities appear across **investment roles**:

    -   Operating Company (OC)

    -   Shareholder (SH)

    -   Controlling Shareholder (CSH)

    -   Global Ultimate Owner (GUO)

-   Summary table: Number of tax haven entities by role. -\> Concentration of tax havens across OC, SH, CSH, and GUO tiers

-   Sankey diagram: Country → Role → Target country. -\> Illustrating investment pathways and the role of tax havens within them

## Impact Assessment
